---
title: "Using Census APIs in R"
author: "Stella Keck, Wei Chang"
date: "7/27/2018"
output:
  prettydoc::html_pretty:
    highlight: github
    theme: architect
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction#
This is a quick tutorial on how to pull census data into R using Census APIs.

###Part 1: Install Required Packages###
Install packages below, if needed use install.packages() to download package first.
```{r}
#construct vector of package names
packages <- c("censusapi","acs","devtools","stringr","plyr","dplyr","stringi", "data.table")

#if packages are absent from library, install them
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}

#load packages
lapply(packages, require, character.only = TRUE)
```

###Part 2: Choose variables, geography, and API###
Access the census variables here: https://api.census.gov/data/2015/acs/acs5/variables.html

Use the same census key below, no need to change:
```{r}
mycensuskey<-"2ca0b2830ae4835905efab6c35f8cd2b3f570a8a"
```

Check census APIs list for name of the API you need to use:
```{r}
availableapis <-listCensusApis()
#View(availableapis)
```

Check different census geographies:
```{r}
availablegeos<-listCensusMetadata(name="acs5",vintage=2015,type="g")
#View(availablegeos)
```

Specify the variables you want to pull from the census API:
```{r}
ACSvars<-c('B17001B_002E', 'B17001B_001E', 'B17001H_002E', 'B17001H_001E', 
           'B17001I_002E', 'B17001I_001E', 'B01001_001E', 'B01001_002E', 
           'B01001_026E', 'B03002_012E', 'B03002_003E', 'B03002_004E', 
           'B03002_006E', 'B01002_001E', 'B17001_001E', 'B17001_002E')
```

###Part 3: Create dataset###
Combine census key with desired API, variables, geography, and region:
```{r}
PUMA_VARS14 <- as.data.table(getCensus(name="acs5",
                        vintage=2014, 
                        key=mycensuskey, 
                        vars=ACSvars,
                        region="public use microdata area:*",
                        regionin="state:36"))

#preview data
head(PUMA_VARS14)
```

(Optional) Manipulate data if desired and add in PUMA column:
```{r}
#Poverty by race
PUMA_VARS14[,blackpovp := B17001B_002E/B17001B_001E]
PUMA_VARS14[,whitepovp := B17001H_002E/B17001H_001E]
PUMA_VARS14[,hisppovp := B17001I_002E/B17001I_001E]

#Black-White poverty ratio
PUMA_VARS14[,BWpovrat := blackpovp/whitepovp]

#Hispanic-White poverty ratio
PUMA_VARS14[,HWpovrat := hisppovp/whitepovp]

#rename columns
setnames(PUMA_VARS14, c("B01001_001E","B01001_002E","B01001_026E","B03002_012E","B03002_003E","B03002_004E","B03002_006E","B01002_001E","B17001_001E","B17001_002E","public_use_microdata_area"), c("totpop","men","women","hisp","nhwhite","nhblack","nhasian","medage","pov12muniv","pov12mlive","puma"))

#Other
PUMA_VARS14[,nhother := (totpop - (hisp + nhwhite + nhblack + nhasian))]
```

Keep desired variables:
```{r}
incvars2 <- c("blackpovp", "whitepovp", "hisppovp", "BWpovrat", 
"HWpovrat", "totpop", "men", "women", "hisp", 
"nhwhite", "nhblack", "nhasian", "medage", "pov12muniv", "pov12mlive", "nhother", "puma")

PUMA_ACS_10_14 <- PUMA_VARS14[,incvars2,with=FALSE]
```

###Part 4: Save data as .csv file or text file###
(Optional) Save .csv file for later use:
```{r}
write.csv(PUMA_ACS_10_14,file="ACSPUMA1014.txt")
```
