---
title: "Fonts in R"
author: "Gretchen Culp, Stella Keck, Wei Chang"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: vignette
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#STEP 1: create a local font directory if one does not already exist#
For more details, see: https://github.com/eosrei/emojione-color-font/issues/14
```{r}
font_dir <- "~/.local/share/fonts"
```

If font directory does not exist, add one:
```{r}
if(!dir.exists(font_dir)){
  system(paste0('mkdir -p ',font_dir))
}
```


#STEP 2: Get font#
Check to see what fonts are already available through R:
```{r}
names(pdfFonts())
```
If your desired font is not already available within R, then download from Google Fonts or another website (.ttf file), extract it from zip format, and place it in your fonts directory you created above. You can do this in R by navigating to the Files tab, clicking the ellipse button and pasting in the address from above. 

This example uses the downloaded file from Google Fonts called "Bungee Inline" - access it here: https://fonts.google.com/?query=bungee+in&selection.family=Bungee+Inline

You can also copy fonts from your workstation which are located in: C/Windows/Fonts


#STEP 3: Install font#
For more information on the system function used below look here: https://stat.ethz.ch/R-manual/R-devel/library/base/html/system.html

You must register the font on the Linux server using the below shell command executed from within R using the system function:
```{r}
system("fc-cache -f -v")
```

To view all available fonts, including fonts from your home directory and fonts available to all users:
```{r}
system("fc-list : family style file spacing")
```
To view fonts only from your home directory, such as your "Bungee Inline" font:
```{r}
system(paste0("fc-list : family style file spacing | grep ",Sys.getenv("USER")))
```


#STEP 4: In order for the font updates to be applied, we must quit the session#
```{r}
.rs.restartR() 
```


#STEP 5: Use your new font#
See simple plot below, insert specified font using the "family" parameter within the text, title, or theme depending on your mapping package.
```{r}
plot(1, pch = 16, cex = 3)
text(1, 1.1, "A fancy dot", family = "Bungee Inline", col = "steelblue", cex = 2)
```

Base Plot example:
```{r}
suppressMessages(library(rgdal))

PUMA.shp <- readOGR(dsn="/opt/shapeFiles/PUMA_2010",layer="PUMA_2010", verbose=FALSE, stringsAsFactors = FALSE)

plot(PUMA.shp, col="white", border="grey", asp=1)
title(main=("NYC PUMA Geography"), family = "Bungee Inline")
```

Ggplot example:
```{r}
suppressMessages(library(broom))
suppressMessages(library(ggplot2))

sug <- readOGR(dsn="/opt/shapeFiles/HRTP",layer="NYC_shp", verbose= FALSE)
sug_f <- tidy(sug)
sug$id <- row.names(sug)
sug_f <- merge(sug_f, sug)

map_theme <- ggplot() +
  geom_polygon(data = sug_f, aes(x = long, y = lat, group = group, fill = value), 
               color = "black", size = 0.25) +
  scale_fill_continuous(name = '% Adults', low = "white", high = "#FF5039", guide = "colorbar", na.value = "grey") +
  labs(title = "% Adults Who Consume Sugary Beverages") +
  coord_equal() +
  theme_void() + theme(plot.title = element_text(hjust = 0.5, family = "Bungee Inline", face = "bold"), legend.title = element_text(family = "Bungee Inline")) #specify font here
print(map_theme)
```


