---
title: "Spatial Join rgeos::over Tutorial"
author: "Stella Keck"
date: "1/24/2019"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction ##
This tutorial covers how to transform a .csv file with longitude and latitude coordinates (in degrees) into a SpatialPointsDataFrame object with the correct NYC projection, which can then be joined spatially with a shapefile (in this example, NTA is the geometry level used). The "over" function assigns the appropriate NTA ID to each point. 

## Tutorial ##

#### Install Required Packages ####
```{r}
suppressMessages(library(sp))
suppressMessages(library(sf))
suppressMessages(library(rgdal))
suppressMessages(library(rgeos))
suppressMessages(library(raster))
```

#### Import dataset and Shapefile ####

```{r}
#replace this .csv with your .csv file#
cb.df <- read.csv("/opt/shapeFiles/CitiBike_stations.csv", stringsAsFactors = FALSE)

#replace this shapefile if using a different geometry besides NTA#
NTA.shp <- readOGR(dsn = "/opt/shapeFiles/NTA_2010_gen", layer = "NTA_2010_gen", stringsAsFactors = F, verbose=F)
```

#### Projections ####
If data is in degrees, change to sf format to manupulate into spatial object and change to feet. Depending on the column names in your dataset, you may need to change "longitude" and "latitude" to the different column names. 
```{r}
#convert data frame to simple features and identify coordinate columns#
cb.sf <- sf::st_as_sf(x = cb.df, coords = c("longitude", "latitude"), crs = 4326, agr = "constant")

#check projection of .shp file#
projection(NTA.shp)

NTA.shp.LISP <- spTransform(NTA.shp, crs("+init=epsg:2263"))

projection(NTA.shp.LISP)
```

#### Convert to sp object ####
```{r}
cb.sp <- as(cb.sf, 'Spatial')
class(cb.sp)

cb.LISP <- spTransform(cb.sp, crs("+init=epsg:2263"))
#projection should now match the .shp file's projection#
projection(cb.LISP)
```


#### Spatial Join ####

```{r}
#find name of NTA ID column to use for spatial join#
head(NTA.shp.LISP@data)

#create new column in points dataset for NTA ID and join#
cb.LISP[["NTA_ID"]] <- over(cb.LISP, NTA.shp.LISP[,"NTA_2010"], fn = NULL)[["NTA_2010"]]

head(cb.LISP)
```

