---
title: "Choropleth Map by PUMA Using Census Poverty Data"
author: "Wei Chang, Stella Keck"
date: "7/20/2018"
output:
  word_document: default
  pdf_document: default
  prettydoc::html_pretty:
    highlight: github
    theme: architect
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Introduction#
This is a quick tutorial for how to make a chlorepleth map using ggplot2 and census data. If not using census data then skip the first section on how to pull census data and go directly to pulling in your shapefile and/or text file (.csv or .txt). 

##Required Packages##

Install packages below, if need use install.packages() to download package first.
```{r}
suppressMessages(library(censusapi))
suppressMessages(library(acs))
suppressMessages(library(devtools))
suppressMessages(library(stringr))
suppressMessages(library(plyr))
suppressMessages(library(dplyr))
suppressMessages(library(stringi))
suppressMessages(library(rgdal))
```

#Pull in Data from Census#
##Choose Variables and Download Census Data Using Census API##
Access the census variables here: https://api.census.gov/data/2016/acs/acs5/variables.html

Check census APIs list for name of the API you need to use:
```{r}
availableapis <-listCensusApis()
#View(availableapis)
```

Check different census geographies that are available:
```{r}
availablegeos<-listCensusMetadata(name="acs5",vintage=2015,type="g")
#View(availablegeos)
```

Specify the variables you want to pull from the census API and assign the census key:
```{r}
ACSpov <- c('GEO_ID', 'PUMA5', 'S1701_C01_001E', 'S1701_C02_001E')

#Use the same census key below, no need to change:
mycensuskey<-"2ca0b2830ae4835905efab6c35f8cd2b3f570a8a"
```

Combine census key with desired API, variables, geography, and region:
```{r}
PUMA_POV <- getCensus(name="acs/acs5/subject",vintage=2016, 
                      key=mycensuskey, vars=ACSpov, 
                      region="public use microdata area:*", regionin="state:36")
head(PUMA_POV)
```

##(Optional) Renaming Columns
```{r}
colnames (PUMA_POV)[colnames(PUMA_POV)=="S1701_C01_001E"]<-"TotalPop"
colnames (PUMA_POV)[colnames(PUMA_POV)=="S1701_C02_001E"]<-"PovPop"
```

##(Optional) Save as an Excel/Text File
```{r}
write.csv(PUMA_POV,file="ACSPoverty1216.txt")
```

#Reading in the PUMA Shapefile#
Download the NYS shapefiles here: https://www.census.gov/geo/maps-data/data/cbf/cbf_puma.html

If producing a NYC map, the NYC shapefiles are available in R:>Data Task Force>Reference_Tips>Geocoding and Mapping>Gretchen's nov 2017 shapefiles
```{r}
NYS_PUMA2016<-readOGR(dsn="/home/health.dohmh.nycnet/skeck/IPUMS/cb_2016_36_puma10_500k",layer="cb_2016_36_puma10_500k",verbose=FALSE)

#check if shapefile was pulled in successfully
plot(NYS_PUMA2016)
```

#Merge Census Data with PUMA Shapefile#
```{r}
#find the variable name of common identifier in both 
str(NYS_PUMA2016@data) #AFFGEOID10
str(PUMA_POV) #GEO_ID

#merge
NYS_m <- merge(NYS_PUMA2016, PUMA_POV, by.x="AFFGEOID10", by.y="GEO_ID", all.x=TRUE)

#check to see if they are merged successfully
str(NYS_m@data)
```

##Calculating Percentage Living Below Poverty Level by PUMA
```{r}
NYS_m$Povpercent <- (NYS_m$PovPop/NYS_m$TotalPop)*100
```

#Mapping Using Ggplot2#
Ggplot2 is a system for 'declaratively' creating graphics, based on "The Grammar of Graphics". You provide the data, tell 'ggplot2' how to map variables to aesthetics, what graphical primitives to use, and it takes care of the details.
```{r}
#Required Packages
suppressMessages(library(broom))
suppressMessages(library(ggplot2))
suppressMessages(library(RColorBrewer))
```
##Fortify the Data##
In order to use ggplot2, the shapefile must be fortified. Make sure you are fortifying the correct merged file (not the sf format merged file for the point data).
```{r}
NYS_f <- tidy(NYS_m)
NYS_m$id <- row.names(NYS_m)
NYS_f <- merge(NYS_f, NYS_m)
head(NYS_f)
```

##Making a Choropleth Map##
```{r}
map1 <- ggplot(data = NYS_f, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = cut_number(Povpercent, 6))) + #specify number of categories
  geom_path(color = "grey", size = 0.25) +
  scale_fill_brewer("% Adults in Poverty", palette = "Spectral") +
  labs(title = "% Adults in Poverty by PUMA") +
  coord_equal() +
  theme_void() + theme(plot.title = element_text(hjust = 0.5, family = "Bookman", face = "bold"))
map1
```

##Making Choropleth Maps with Customized Number of Categories##

###Four Categories###
```{r}
#create a color function
colfunc <- colorRampPalette(c(hsv(0.06,0.25,1),hsv(0.06,1,1)))

#cut data into 4 categories#
my.cols.2 <- colfunc(4) #specify number of colors in palette
NYS_f$cutpov2<- 
  cut(NYS_f$Povpercent, breaks = c(0, 10, 20, 30, 100),
      labels = my.cols.2,
      include.lowest = TRUE)

NYS_f$cutpov3<- 
  cut(NYS_f$Povpercent, breaks = c(0, 10, 20, 30, 100),
      include.lowest = TRUE)
```

```{r}
#plot by cut (4)#
mapcut_4 <- ggplot(data = NYS_f, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = cutpov2)) + 
  geom_path(color = "grey48", size = 0.25) +
  scale_fill_brewer("% Adults in Poverty", palette = "PuRd", #substitute"PuRd" with other palette names from the "Optional: "Choosing Color Palettes" step below
                    labels = c("0-10", "10-20", "20-30", "30+")) +
  labs(title = "% Adults in Poverty by PUMA") +
  coord_equal() +
  theme_void() + theme(plot.title = element_text(hjust = 0.5, family = "Bookman", face = "bold"))
mapcut_4


mapcut_10 <- ggplot(data = NYS_f, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = cutpov3)) + 
  geom_path(color = "grey48", size = 0.25) +
  scale_fill_brewer("% Adults in Poverty", palette = "PuRd", #substitute"PuRd" with other palette names from the "Optional: "Choosing Color Palettes" step below
                    labels = c("0-10", "10-20", "20-30", "30+")) +
  labs(title = "% Adults in Poverty by PUMA") +
  coord_equal() +
  theme_void() + theme(plot.title = element_text(hjust = 0.5, family = "Bookman", face = "bold"))
mapcut_10
```

###Six Categories###
```{r}
#cut data into 6 categories#
my.cols <- colfunc(6) #specify number of colors in palette
NYS_f$cutpov<- 
  cut(NYS_f$Povpercent, breaks = c(0, 5, 10, 20, 30, 40, 100),
      labels = my.cols,
      include.lowest = TRUE)
```

```{r}
#plot by cut (6)#
mapcut_6 <- ggplot(data = NYS_f, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = cutpov)) + 
  geom_path(color = "grey48", size = 0.25) +
  scale_fill_brewer("% Adults in Poverty", palette = "PuRd", #substitute"PuRd" with other palette names from the "Optional: "Choosing Color Palettes" step below
                    labels = c("0-5", "5-10","10-20", "20-30", "30-40", "40+")) +
  labs(title = "% Adults in Poverty by PUMA") +
  coord_equal() +
  theme_void() + theme(plot.title = element_text(hjust = 0.5, family = "Bookman", face = "bold"))
mapcut_6
```

###(Optional) Choosing Color Palettes###
```{r}
library (RColorBrewer)
display.brewer.all(4)#specify the number of categories to preview palette
```